Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-01 paper/paper.tex
impact of negative examples
pass

%\vspace{-1ex}
\begin{table}[t!]
  \footnotesize
  \caption{\label{table:rules}Example of capture-relevant options of Suricata.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{clp{6cm}}
    \toprule
    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
    \midrule     
    1 & dsize & matches on the size of the packet payload\\
    2 & itype &  matches on a specific ICMP type\\
    3 & icode & matches on a specific ICMP code\\
    4 & icmp\_seq  & checks a ICMP sequence number\\
    5 & icmp\_id & matches on specific ICMP id-values\\
    6 & window & checks the size of the TCP window\\
    7 & flags & matches on TCP flags\\
    8 & fragbits & checks if the frag. or reserved bits are set in the IP header\\
    9 & threshold & minimum threshold for a rule before it generates alerts\\
    10 & content & checks if argument is present on the payload\\
    \bottomrule
  \end{tabular}
  \vspace{-3ex}  
\end{table}

@@

\setlength{\tabcolsep}{8pt}
\begin{table}[h!]
  \footnotesize
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
    2100384 & 2 & 18$\uparrow$& 2 & 2 & 2 \\
    2100399 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    2100401 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100402 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100408 & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
    2100485 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
    2100486 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    %%UDP Flood & - & - & - & - & - \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 4 & 4 & 2\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
    JBoss JMX & 87 & 87 & 15\cellcolor{lightgray}$\downarrow$ & 84 & 269$\uparrow$ \\
    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
    WeBid LFI & 17 & 17 & 144$\uparrow$ & 27$\uparrow$ & 41$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
    .idq access & 2  & 2 & 1 & 1 & 15$\uparrow$ \\
    /system32/ in URI & 2  & 2 & 6$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$\\
    Basebuilder LFI & 48  & 48 & 846$\uparrow$ & 65$\uparrow$ & 664$\uparrow$\\
    Tomcat Snoop Access & 37  & 37 & 1797$\uparrow$ & 6\cellcolor{lightgray}$\downarrow$ & 3580$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

\renewcommand{\arraystretch}{0.9}
\begin{table*}[t!]
  \footnotesize
  \caption{\label{table:attacks}-}
  \vspace{-2ex}
  \centering
    \begin{tabular}{lp{2cm}p{2.1cm}ccp{2cm}c}
    \toprule
    \multicolumn{1}{c}{Paper} &
    \multicolumn{1}{c}{Input} &
    \multicolumn{1}{c}{Dataset} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Output} &
    \multicolumn{1}{c}{Evaluation} &
    \multicolumn{1}{c}{Availability} \\
    \midrule
    \cite{vollmer-etal-cics2011} & Malign Traffic & Manually crafted & ICMP & Rule List & TP, FP & No\\
    \midrule
    \cite{fallahi} & Mixed Traffic & \cite{iscx} & HTTP, SSH, TCP & Single Rule & TP, FP, Precision, Recall, F1 & No\\    
    \midrule
    \cite{Kao2015AutomaticNR} & Malign Traffic, Benign Traffic, Rule Dataset & Manually crafted, \cite{emerging-threats-open} & HTTP & Single Rule & Generating time, FP & No \\
    \midrule
    \cite{guruprasad}& Malign Traffic & \cite{iscx} \cite{darpa1999} & ICMP, TCP & Rule List & FP & No\\
    \midrule
    Syrius & Malign Traffic, Benign Traffic, Rule Dataset & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Rule Ranking & Yes \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

This section evaluates sensitivity of \tname{} to the number of
attacks provided on input. Recall that \tname\ uses one example attack
to generate the seed rule (at the Reverse Engineering step) and the
remaining attacks to evaluate soundness of the rule (at the Rules
Creations step). It is expected that the number and variety of
attacks---of the same kind---plays an important role in the
performance of the technique. To answer this research question, we selected two cases with the
highest number of examples (see column ``\#~Vars'' on
Table~\ref{table:results})---idqaccess and scripttag and measured the
position of the golden rule at the rank for an increasing number of
examples provided to \tname. For each number of variants, we sampled
five sets of that size to provide as input to the
technique. Figure~\ref{fig:impact-number-attacks} shows the results
for each attack we selected. The y-axis of each plot denotes the rank
of the golden rule and x-axis denotes the number of attack examples
used to synthesize rules. Each point in the x-axis is associated with
the boxplot of the distribution of results for the corresponding
number of variants. Results indicate that 1) the number of examples is
important, 2) the choice of examples is important, and 3) a small
number of examples suffices to reach the best results.

\begin{figure}[t!]
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \caption{scripttag}    
  \end{subfigure}%
\vspace{2ex}  
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
    \caption{coldfusion}
  \end{subfigure}%
    \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \caption{inc}    
  \end{subfigure}%
    \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \caption{jsp}    
  \end{subfigure}%

%\vspace{-2.5ex}  
\caption{Impact of number of examples.}
\label{fig:impact-number-attacks}
%\vspace{1ex}
\end{figure}

\begin{center}
\begin{tcolorbox}[enhanced,width=3.4in,center upper,drop shadow southwest,sharp corners]
As expected, \tname\ is sensitive to the number and choice of example
input attacks. However, a small number of examples suffices to achieve
best results.
\end{tcolorbox}
\end{center}

@@

\subsubsection{Lack of comparison with other tools}
 \Gui{There are a few reasons that makes the comparison between Syrius and other tools impossible

\textbf{Tools availability.} We tried contact with authors of 4 different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}\cite{Kao2015AutomaticNR}, but got no response from them. Since their tools are not available online or at least not publicly, we could not run them.

\textbf{Different output and specific rules.} We considered using the results they provides with results that Syrius provide, but concluded it wouldn't be possible as the results are essentially different. Commonly the  other results focus on showing the ratio of false negatives and false positives as a way to show the efficiency of the learning algorithm, but never comparing the generated rules to a Ground Truth, and most of times not showing individually the attacks used. Syrius' results, on the other hand, shows the generated rules that produce no false positives and negatives to the given packets, and always compares the results with a Ground Truth to verify its efficiency.

\textbf{'content' option support.} Another limitation is the support of 'content' option. Most papers do not support rule generation with the 'content' option, using only classic DDoS attacks in the rule generation, like ICMP and SYN flood, which only use numerical and categorical options. Some papers do support 'content', but don't explain well how they handle those rules with exclusive 'content'.}
%%  \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 
%%  evolutionaries approaches to 
%% autonomously generate rules using the Snort IDS. They built 
%% their own ICMP datasets by using packet creation tools or by capturing network traffic. 
%% Guruprasad also used DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets 
%% for non-ICMP attacks.
%% 
%% Vollmer created ten different packets to match ten existing rules, and then used 
%% these packets as input to the rule generation process. As a result, the produced 
%% rules achieved a 100\% detection rate and produced 3 false positives out of 33k packets.
%% 
%% Guruprasad manually created a set of rules by analyzing the network patterns and 
%% the datasets, this set of rules was used as the initial population of the algorithm. 
%% The produced rules achieved a detection rate up to 96\%, the false positive rate 
%% was not evaluated. 
%% 
%% On our experiments with Syrius, we used a total of \totalPackets packets that 
%% matches \totalAttacks different existing rules and it was able to generate rules that detected 
%% all the packets with no false positives.  
%%  }  
\subsubsection{Running time}\Mar{}\Gui{We ran all our tests using an Intel(R) Xeon(R) CPU E3-1240 v6 and 8GB RAM. For attacks with only categorical and numerical options, the average running time was 6 seconds per attack. For attacks with the content option, the average running time was 1445 seconds.}

>>>>

%\vspace{-1ex}
\begin{table}[t!]
  \footnotesize
  \caption{\label{table:rules}Example of capture-relevant options of Suricata.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{clp{6cm}}
    \toprule
    \multicolumn{1}{c}{\#} & \multicolumn{1}{c}{Name} &  \multicolumn{1}{c}{Description}\\
    \midrule     
    1 & dsize & matches on the size of the packet payload\\
    2 & itype &  matches on a specific ICMP type\\
    3 & icode & matches on a specific ICMP code\\
    4 & icmp\_seq  & checks a ICMP sequence number\\
    5 & icmp\_id & matches on specific ICMP id-values\\
    6 & window & checks the size of the TCP window\\
    7 & flags & matches on TCP flags\\
    8 & fragbits & checks if the frag. bits are set in the IP header\\ % or reserved bits
    9 & threshold & min \# messages for a rule before it generates alerts\\
    10 & content & checks if argument is present on the payload\\
    \bottomrule
  \end{tabular}
  \vspace{-3ex}  
\end{table}

@@

\begin{table}[h!]
  \footnotesize
  \setlength{\tabcolsep}{8pt}  
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
    2100384 & 2 & 18$\uparrow$& 2 & 2 & 2 \\
    2100399 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    2100401 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100402 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100408 & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
    2100485 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
    2100486 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    %%UDP Flood & - & - & - & - & - \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 4 & 4 & 2\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
    JBoss JMX & 87 & 87 & 15\cellcolor{lightgray}$\downarrow$ & 84 & 269$\uparrow$ \\
    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
    WeBid LFI & 17 & 17 & 144$\uparrow$ & 27$\uparrow$ & 41$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
    .idq access & 2  & 2 & 1 & 1 & 15$\uparrow$ \\
    /system32/ in URI & 2  & 2 & 6$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$\\
    Basebuilder LFI & 48  & 48 & 846$\uparrow$ & 65$\uparrow$ & 664$\uparrow$\\
    Tomcat Snoop Access & 37  & 37 & 1797$\uparrow$ & 6\cellcolor{lightgray}$\downarrow$ & 3580$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table*}[t!]
  \footnotesize
  \setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:comparison}Alternative techniques.}
  \vspace{-2ex}
  \centering
    \begin{tabular}{lcccccc}
    \toprule
    \multicolumn{1}{c}{Paper} &
    \multicolumn{1}{c}{Input} &
    \multicolumn{1}{c}{Output} &    
    \multicolumn{1}{c}{Dataset} &
    \multicolumn{1}{c}{Protocols} &
    \multicolumn{1}{c}{Evaluation} &
    \multicolumn{1}{c}{Avail.?} \\
    \midrule
    \cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
%    \midrule
    \cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
%    \midrule
    \cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
    One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
%    \midrule
    \cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
    ICMP, TCP & FP & No\\
%    \midrule
    Syrius & Malign/Benign Traffic, Rule Dataset & Ranking & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Yes \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Creations step). It is expected that the number and variety of attacks
play a role in the performance of the technique.

\begin{figure}[h]
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}       \caption{scripttag}    
  \end{subfigure}%
%\vspace{1ex}  
\caption{Impact of number of positive examples.}
\label{fig:impact-number-attacks}
%\vspace{1ex}
\end{figure}

We answer this question in two parts. Firts, we analyzed the impact of
the amount of positive examples in results. For that, we selected two
cases with the highest number of examples (see column ``\#~Vars'' on
Table~\ref{table:results})---idqaccess and scripttag and measured the
position of the golden rule at the rank for an increasing number of
examples provided to \tname. For each number of variants, we sampled
five sets of that size to provide as input to the
technique. Figure~\ref{fig:impact-number-attacks} shows the results
for each attack we selected. The y-axis of each plot denotes the rank
of the golden rule and x-axis denotes the number of attack examples
used to synthesize rules. Each point in the x-axis is associated with
the boxplot of the distribution of results for the corresponding
number of variants. Results indicate that 1) the number of examples is
important, 2) the choice of examples is important, and 3) a small
number of examples suffices to reach the best results.

\begin{figure}[h!]
 \centering
 \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
    \caption{coldfusion}
  \end{subfigure}%
    \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \caption{inc}    
  \end{subfigure}%
    \hspace{-8ex}
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \caption{jsp}    
  \end{subfigure}%
%\vspace{-2.5ex}  
\caption{Impact of number of negative examples.}
\label{fig:impact-negative}
\vspace{1ex}
\end{figure}

We conducted a similar experiment with negative examples on three
different attacks. Figure~\ref{fig:impact-negative} shows
results. Note that, compared with the plots of
Figure~\ref{fig:impact-number-attacks}, the rank of the golden rule
improves at a much slower pace. Fortunately, negative traffic is much
easier to access.

\begin{center}
\begin{tcolorbox}[enhanced,width=3.5in,center upper,drop shadow southwest,sharp corners]
\tname\ is sensitive to the number and choice of examples. However, a
small number of positive examples suffices to achieve good results.
\end{tcolorbox}
\end{center}

@@

\subsubsection{Lack of comparison with other tools}
Table~\ref{table:comparison} summarizes the characteristics of
alternative techniques and provides details about the evaluation used
in their original paper. Unfortunately, side comparison was not
possible for three key reasons. First, we did not have access to the
tools. We tried, unsuccssefully, making contact with authors to obtain
the tools. Second, the tools output different objects. For that
reason, a comparison based on published results did not make sense. It
is important to recall that we did not report precision and recall as
it is 100\% always. \tname captures, by construction, all positive
traffic and does not capture any negative traffic. We used a stronger
criteria instead---the position of the correct rule in the ranking. It
is possible that other rules, higher in the ranking, are useful, but
only domain experts can identify that. Third, we are the only
technique to support content options, which are prevalent and
challenging to handle. A side comparison would be unfair if we
consider rules with those options.

%%  \Gui{There are a few reasons that makes the comparison between Syrius and other tools impossible

%% \textbf{Tools availability.} We tried contact with authors of 4 different papers \cite{vollmer-etal-cics2011}\cite{muthuregunathan-etal2009}\cite{fallahi}\cite{Kao2015AutomaticNR}, but got no response from them. Since their tools are not available online or at least not publicly, we could not run them.

%% \textbf{Different output and specific rules.} We considered using the results they provides with results that Syrius provide, but concluded it wouldn't be possible as the results are essentially different. Commonly the  other results focus on showing the ratio of false negatives and false positives as a way to show the efficiency of the learning algorithm, but never comparing the generated rules to a Ground Truth, and most of times not showing individually the attacks used. Syrius' results, on the other hand, shows the generated rules that produce no false positives and negatives to the given packets, and always compares the results with a Ground Truth to verify its efficiency.

%% \textbf{'content' option support.} Another limitation is the support of 'content' option. Most papers do not support rule generation with the 'content' option, using only classic DDoS attacks in the rule generation, like ICMP and SYN flood, which only use numerical and categorical options. Some papers do support 'content', but don't explain well how they handle those rules with exclusive 'content'.}
%%  \Luc{Vollmer \cite{vollmer-etal-cics2011} and Guruprasad \cite{guruprasad} used 
%%  evolutionaries approaches to 
%% autonomously generate rules using the Snort IDS. They built 
%% their own ICMP datasets by using packet creation tools or by capturing network traffic. 
%% Guruprasad also used DARPA 1999 \cite{darpa1999} and ISCX 2012 \cite{iscx} datasets 
%% for non-ICMP attacks.
%% 
%% Vollmer created ten different packets to match ten existing rules, and then used 
%% these packets as input to the rule generation process. As a result, the produced 
%% rules achieved a 100\% detection rate and produced 3 false positives out of 33k packets.
%% 
%% Guruprasad manually created a set of rules by analyzing the network patterns and 
%% the datasets, this set of rules was used as the initial population of the algorithm. 
%% The produced rules achieved a detection rate up to 96\%, the false positive rate 
%% was not evaluated. 
%% 
%% On our experiments with Syrius, we used a total of \totalPackets packets that 
%% matches \totalAttacks different existing rules and it was able to generate rules that detected 
%% all the packets with no false positives.  
%%  }  
\subsubsection{Running time}\Mar{}\Gui{We ran all our tests using an Intel(R) Xeon(R) CPU E3-1240 v6 and 8GB RAM. For attacks with only categorical and numerical options, the average running time was 6 seconds per attack. For attacks with the content option, the average running time was 1445 seconds.}
